---
title       : Fuel Consumption
subtitle    : Calculates average fuel consumption based on numbers of cylinders and gears
author      : Sasa Bogdanovic
job         : Chief Data Analyst at Ruby Industries Inc.
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction

Fuel Consumption is Shiny application that calculates average fuel consumption in miles per gallon, depending on numbers of cylinders and gears.

It can be accessed at https://sasha.shinyapps.io/cars.

Data used for calculations is mtcars dataset.

Users can easily change numbers of cylinders and gears using slider and radio butttons options.

--- .class #id 

## Application Layout - sidebar panel

Layout is organized as a page with sidebar and main panel.

Sidebar contains:
- usage instructions - basic instructions to help user understand the purpose of the application, and how to use input controls to change the output.
- input controls 
  - slider control for numbers of cylinders
  - radio buttons for number of gears

---

## Application layout - main panel
Main panel is organized as a tab set with the following tabs:
- MPG vs Number of Cylinders - contains reactive scatterplot with MPG ~ #Cylinders points, and overlayed lines for average full dataset mpg (grey color) and average mpg for selected numbers of cylinders and gears (red color).
- MPG vs Number of Gears - contains reactive scatterplot with MPG ~ #Gears points, and overlayed lines for average full dataset mpg (grey color) and average mpg for selected numbers of cylinders and gears (red color).
- Data table - contains reactive data table filtered for selected numbers of cylinders and gears.

Main panel also shows
- reactive text with average MPG for currently selected options (updated with every change).
- static text with information on combinations of numbers of cylinders and gears that yield minimum nad maximum fuel consumption.

---

## Sample plots and data table

```{r echo=FALSE}
library(shiny)
library(doBy)

data(mtcars)
mn <- mean(mtcars$mpg)

mtcarsSub <- subset(mtcars, cyl == 8 & gear == 3)
mns <- mean(mtcarsSub$mpg)

mnv <- summaryBy(mpg ~ cyl + gear, data=mtcars, FUN=c(mean))
optMin <- subset(mnv, mpg.mean == min(mnv$mpg.mean))
optMax <- subset(mnv, mpg.mean == max(mnv$mpg.mean))
```

Sample code and plots

```{r fig.height = 4, fig.align = 'center'}
par(mfrow=c(1,2))
plot(mtcars$cyl, mtcars$mpg, col="blue", xlab = "Number of cylinders", ylab = "Miles per gallon", main="MPG vs #Cyl")
lines(c(4, 8), c(mn, mn), col="darkgrey", lwd=2); lines(c(4, 8), c(mns, mns), col="red", lwd=2)
plot(mtcars$gear, mtcars$mpg, col="blue", xlab = "Number of gears", ylab = "Miles per gallon", main="MPG vs #Gears")
lines(c(3, 5), c(mn, mn), col="darkgrey", lwd=2); lines(c(3, 5), c(mns, mns), col="red", lwd=2)
```
---
